Crystal as a thermodynamic computer

### 2024-09-14 paper/main.tex
update

To overcome the overlap gap property, we propose to do block annealing, by introducing a heat zone with width $W$.
It represents the region of the system that interacts with the heat bath, as shown in \Cref{Surface-programmable-thermodynamic computer}.
Let layer $i$ be defined as the set of spins with x-coordinate $i$.
The system is devided into $\lceil\frac{M}{W}\rceil$ blocks,
where block $j$ consists of layers in the range $[(j-1)W+1, jW]$.
The system can then be cooled by sequentially cooling the blocks from left to right with an heat bath, where the temperature of the heat bath is gradually decreased.
We treat it as an ideal bath as it maintains a space-invariant temperature within the heat zone while keeping other regions at zero temperature.

Assume the SA procedure in each block has an error probability $\epsilon$.
Repeating the procedure in $\frac{M}{W}$ blocks results in a success probability of $(1 - \epsilon)^{\frac{M}{W}}$ for finding the global ground state. 
To keep this probability constant as $M \to \infty$, the error probability $\epsilon$ should scale as $\frac{W}{M}$.

Now we turn to determining the sweep time $t_{th}$ required to achieve $\epsilon$ error in each procedure. 
Let $\lambda_1$ and $\lambda_2$ be the largest and second largest eigenvalues of the Markov transition matrix, respectively. 
We assume the relationship between $\frac{\lambda_2}{\lambda_1}$ and $W$ can be expressed as
\begin{equation}
    \label{eq:spectral-gap}
    1 - \frac{\lambda_2}{\lambda_1} \sim e^{-W}.
\end{equation}
Consequently, $t_{th}$ is related to $\epsilon$ by:
\begin{equation}
    \left( \frac{\lambda_2}{\lambda_1} \right) ^ {t_{th}} = \epsilon \rightarrow t_{th} \sim \ln(\frac{M}{W})e^{W},
\end{equation}
and the total sweep times scale as $\frac{M}{W}\ln(\frac{M}{W})e^W$.

@@

In practice, the heat bath and heat flux should establish a temperature gradient in the system.
% Suppose the heat bath only connects to a small rectangle region of the system.
The temperature $T(x, t)$  at a position with x-coordinate  $x$  and time  $t$  can be modeled as  $T(x, t) \propto \Lambda^{-|x - L(t)|}$ , 
where $\Lambda$ is a heat flux constant and $L(t)$ denotes the center of the heat bath at time $t$.

The block annealing process then can be reconstruct as moving
the heat bath from the left surface to the right surface at speed $v$. Defining $\Delta E_{max}$ to be the
largest possible energy change when a spin is fliped.
The effective heat zone with width $W'(\Lambda)$ should then be $[y-W'(\Lambda), y]$, where position $y$ satisfying 
$\exp(\frac{\Delta E_{max}}{T(y)}) = 0.1$ and $y<L(t)$. 
% Since the layers in fronts of the effective heat zone can only be fliped with a $\epsilon$ probability if the local energy would increase. 
The effective width $W'$ then should scale as $\exp(-\Delta E_{max} \Lambda^{cW'(\Lambda)}) \sim \epsilon$, where $c$ is an empirical constant and can be determined during the experiment. 
\ym{I am not sure about how to give a concrete definition of the effective width $W'$.}

@@

\begin{equation}
    \label{eq:total-annealing-time}
    t_{total} \propto \frac{M}{W'(\Lambda)}(\ln(\frac{M}{W'(\Lambda)}))^{1+ \frac{1}{c\ln(\Lambda)}}
\end{equation}

To determine the relationship between the total annealing time and the input size $N$, we should
further discuss the relationship between the spectral gap given in  \Cref{eq:spectral-gap} and inputs size $N$.
However, we failed to find a direct relationship.
Numerical result shows that computing complexity is linear to $N$. Thus a total sweep time
for the computation should scale as $NM\log^{1 + \frac{1}{c\ln(\Lambda)}}(M)$.
\ym{Please help me to think about how to evaluate the time complexity while $N$ increases.} 
The moving speed $v$ can be then calculated as

>>>>

To overcome the overlap gap property, we propose to do block-wise annealing.
As shown in \Cref{Surface-programmable-thermodynamic computer}, we divide the lattice into $\frac{M}{W}$ blocks along the time dimension, each having width $W$.
%by introducing a heat zone with width $W$.
%It represents the region of the system that interacts with the heat bath, as shown in \Cref{Surface-programmable-thermodynamic computer}.
%Let layer $i$ be defined as the set of spins with x-coordinate $i$.
These blocks are initialized to a low enough temperature, so that the spins freeze into a random local minimum.
Then we anneal these blocks sequentially from left to right by heating up a block to a temperature higher than the energy scale and then cooling it down.
%We treat it as an ideal bath as it maintains a space-invariant temperature within the heat zone while keeping other regions at zero temperature.
Given a slow enough cooling rate, the spins in the block will settle into the global ground state.
The cooling rate determines the true computing speed of this model.
It must scale with the number of computational steps $M$, otherwise, the probability of success will exponentially decrease with the number of blocks.
How cooling rate should scale with $M$ becomes a key question in this work.

Assume the SA procedure in each block has an error probability $\epsilon$.
Repeating the procedure in $\frac{M}{W}$ blocks results in a success probability of $(1 - \epsilon)^{\frac{M}{W}}$ for finding the global ground state. 
To fix the success probability for arbitrary $M$, we must have $\epsilon \sim \frac{W}{M}$.
The annealing time $t_{th}$ required to achieve a certain error in each block is determined by the spectral gap $\Delta$ of the Markov transition matrix of the SA update on the block.
%Since the largest eigenvalue of a transition matrix is 1, the spectral gap is $\Delta = 1 - \lambda_2$, where $\lambda_2$ is the second-largest eigenvalue.
The scaling behavior of $\Delta$ and block size $W$ can be complicated, in the worst case, it can be exponential, i.e.,
\begin{equation}
    \label{eq:spectral-gap}
    \Delta \sim e^{-W}.
\end{equation}
Consequently, $t_{th}$ is related to $\epsilon$ by $\left( 1-\Delta \right) ^ {t_{th}} = \epsilon$, which gives $t_{th} \sim \ln(\frac{M}{W})e^{W}$, and the total sweep times scale as $\frac{M}{W}\ln(\frac{M}{W})e^W$.

@@

%In practice, the heat bath and heat flux should establish a temperature gradient in the system.
% Suppose the heat bath only connects to a small rectangle region of the system.
Instead of block-wise annealing, we propose to move the heat bath from the left surface to the right surface at a constant speed $v$.
At time $t$, the temperature distribution $T$ is given by $T_{\rm max}\Lambda^{-|x - vt|}$, where $x$ is the computational step, $T_{\rm max}$ is the maximum temperature and $\Lambda > 1$ is a constant that determines the width of the heat zone.
%The temperature $T(x, t)$  at a position with x-coordinate  $x$  and time  $t$  can be modeled as  $T(x, t) \propto \Lambda^{-|x - L(t)|}$ , 
%where $\Lambda$ is a heat flux constant and $L(t)$ denotes the center of the heat bath at time $t$.

We can define an effective heat zone with width $W'$ as the width of the raising edge of the temperature distribution, where the spins can be cooled actively.
%Defining $\Delta E_{max}$ to be the largest possible energy change when a spin is fliped.
%The effective heat zone with width $W'(\Lambda)$ should then be $[y-W'(\Lambda), y]$, where position $y$ satisfying 
%$\exp(\frac{\Delta E_{max}}{T(y)}) = 0.1$ and $y<L(t)$. 
% Since the layers in fronts of the effective heat zone can only be fliped with a $\epsilon$ probability if the local energy would increase. 
The spins outside the effective heat zone are either frozen (too cold) or flip randomly (too hot).
As a spin flips with a probability $\sim e^{-\Delta E/T}$, the cold edge of the heat zone can be defined as the position where the probability of flipping is $\epsilon$.
We have $W'$ scales with the required error probability $\epsilon$ as:
\begin{equation}
    e^{-\frac{\Delta E}{T_{\rm max}}\Lambda^{cW'}} \sim \epsilon
\end{equation}
where $c$ is an empirical constant and can be determined during the experiment. 
\ym{I am not sure about how to give a concrete definition of the effective width $W'$.}\jinguo{How about this?}

@@

\begin{equation}
    \label{eq:total-annealing-time}
    t_{total} \propto \frac{M}{W'}(\ln(\frac{M}{W'}))^{1+ \frac{1}{c\ln(\Lambda)}}
\end{equation}

To determine the relationship between the total annealing time and the input size $N$, we should
further discuss the relationship between the spectral gap given in  \Cref{eq:spectral-gap} and inputs size $N$.
However, we failed to find a direct relationship.
Numerical result shows that computing complexity is linear to $N$. Thus a total sweep time
for the computation should scale as $NM\log^{1 + \frac{1}{c\ln(\Lambda)}}(M)$.
\ym{Please help me to think about how to evaluate the time complexity while $N$ increases.}\jinguo{Leave it to Maddie}
The moving speed $v$ can be then calculated as
